# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
```{r}
unzip("activity.zip")

dat =read.csv("activity.csv")
dat$date <- as.Date(dat$date)

stepsByDay = aggregate(steps ~date, data=dat, sum)
stepsByInt = aggregate(steps ~interval, data=dat, mean)
```



## What is mean total number of steps taken per day?
```{r}
hist(stepsByDay$steps, main="Mean number of steps per day")
```

```{r}
mns = mean(stepsByDay$steps, na.rm = TRUE)
mds = median(stepsByDay$steps, na.rm = TRUE)
```
The mean total number of steps per day is `r mns`; the median number of steps is `r mds`.

## What is the average daily activity pattern?

1. Time series plot
```{r }
maxInt= stepsByInt[which(max(stepsByInt[,2])==stepsByInt[,2]),]
plot(steps~interval, dat=stepsByInt, type="l")
abline(v=maxInt[1], col="red")
```

2. Which 5-min interval contains the maximum number of steps?


The maximum number of steps is taken during the interval # `r maxInt[1]`.

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
nas = table(is.na(dat$steps))[["TRUE"]]
```
The number of rows with missing data is `r nas`.

2. Devise a strategy
Use mean interval value for missing data.

3. Create a new dataset with missing data filled in.
```{r}
naInt= dat[which(is.na(dat$steps)),]

# Get corresponding interval values for
xx = merge(naInt, stepsByInt, by="interval")
#naInt$steps = xx$steps.y
rn=as.integer(row.names(naInt))
dat[rn,"steps"] = xx$steps.y
```
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}
stepsByDay = aggregate(steps ~date, data=dat, sum)
mns2 = mean(stepsByDay$steps)
mds2 = median(stepsByDay$steps)

```

```{r}
hist(stepsByDay$steps, main="Mean number of steps per day")
```

The mean of steps after filling in missing data is `r mns2`; the median is `r mds2`. The mean remains the same, the median is a bit higher.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
wd=weekdays(dat$date)
typeOfDay = wd == "Sunday" | wd=="Saturday"
typeOfDay[typeOfDay==TRUE] <- "weekend"
typeOfDay[typeOfDay==FALSE] <- "weekday" 
typeOfDay <- as.factor(typeOfDay)
dat = cbind(dat, typeOfDay)
```


```{r}
stepsByIntWE = aggregate(steps ~interval, data=dat[dat$typeOfDay=="weekend",], mean)

stepsByIntWD = aggregate(steps ~interval, data=dat[dat$typeOfDay=="weekday",], mean)

par(mfrow=c(2,1))
plot(steps~interval, dat=stepsByIntWE, type="l", main="weekend", col="blue")
plot(steps~interval, dat=stepsByIntWD, type="l", main="weekdays", col="blue")

```


